---
output:
  html_document: default
---

# Summary: NOB1-heat-shock- RT-qPCR experiment_draft.

## Setup knitr options and load packages

```{r setup,warning=FALSE,message=FALSE}

# knitr options for report generation
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE,
  results = "show")

# Load packages
library(tidyr)
library(ggplot2)
library(dplyr)
library(readr)
library(tidyqpcr)
library(stringr)

# set default theme for graphics
theme_set(theme_bw(base_size = 11) %+replace%
  theme(
    strip.background = element_blank(),
    panel.grid = element_blank()
  ))
```

## Organize sample and target data on the plates

```{r set_platedata_NOB1_1,dependson="label_plates"}
# plate_1
# set unexclude wells for NOB1
letters_series <- rep(LETTERS[5:16], each = 24)
numbers_series <- rep(1:24, times = 3)
wells_exclude <- paste0(letters_series, numbers_series)

plate_cq_extdata_1_2nd_derivative <- read_lightcycler_1colour_cq("WD_250723_NOB1_HS_2nd_derivative.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

# set plate distribution

plate_cq_data_1_2nd_derivative <- 
  create_blank_plate(well_row = LETTERS[1:4], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_1_2nd_derivative, by = "well") # %>% # join that info to loaded data

target_id_levels <- c("NOB1","PGK1","RPS3","RPS13")
rowkey4 <- tibble(
  well_row = LETTERS[1:4],
  target_id = target_id_levels)
sample_id_levels <- c("NOB1swap_nonHS_rep1", "WT_nonHS_rep2", "NOB1swap_nonHS_rep2", "WT_nonHS_rep1", "NOB1swap_shortHS_rep1", "NOB1swap_shortHS_rep2")
prep_type_values <- c(rep("+RT", times = 18), rep("-RT", times = 6))
colkey12 <- tibble(
  well_col = 1:24,
  sample_id = rep(sample_id_levels, times = 4),
  prep_type = prep_type_values)

plate_cq_data_1_2nd_derivative <- label_plate_rowcol(
  plate = plate_cq_data_1_2nd_derivative,
  rowkey = rowkey4,
  colkey = colkey12)

# plate_2
# set unexclude wells for ARG8
letters_series <- c(rep(LETTERS[1:4], each = 24), rep(LETTERS[9:16], each = 24))
numbers_series <- rep(1:24, times = 3)
wells_exclude <- paste0(letters_series, numbers_series)

plate_cq_extdata_2_2nd_derivative <- read_lightcycler_1colour_cq("WD_250723_NOB1_HS_2nd_derivative.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

# set plate distribution

plate_cq_data_2_2nd_derivative <- 
  create_blank_plate(well_row = LETTERS[5:8], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_2_2nd_derivative, by = "well") # %>% # join that info to loaded data

target_id_levels <- c("NOB1","PGK1","RPS3","RPS13")
rowkey4 <- tibble(
  well_row = LETTERS[5:8],
  target_id = target_id_levels)
sample_id_levels <- c("WT_shortHS_rep1", "WT_shortHS_rep2", "NOB1swap_mediumHS_rep1", "NOB1swap_mediumHS_rep2", "WT_mediumHS_rep1", "WT_mediumHS_rep2")
prep_type_values <- c(rep("+RT", times = 18), rep("-RT", times = 6))
colkey12 <- tibble(
  well_col = 1:24,
  sample_id = rep(sample_id_levels, times = 4),
  prep_type = prep_type_values)

plate_cq_data_2_2nd_derivative <- label_plate_rowcol(
  plate = plate_cq_data_2_2nd_derivative,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_NOB1_2nd_derivative = rbind(plate_cq_data_1_2nd_derivative, plate_cq_data_2_2nd_derivative)
plate_cq_data_NOB1_2nd_derivative

```


## Plot unnormalized Cq data

```{r plot_unnormalized_NOB1,dependson="load_plates",fig.height=6,fig.width=8}
unnormalized_plot_data <-  plate_cq_data_NOB1_2nd_derivative %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))
unnormalized_plot_data

# 2nd derivative approach
strain_colors <- c("#330099","#3366FF","#66CCFF","#FF0000","#FF3399","#FF99CC")
ggplot(data = unnormalized_plot_data) +
  geom_point(aes(x = target_id, y = cq, shape = rep_number, colour = clone),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "Unnormalized_NOB1_2nd_derivative_cq"
  ) +
  facet_wrap(~prep_type,ncol=5) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)
```


# Normalized data: delta Cq

## Normalize Cq to PGK1, RPS3 and RPS13, within Sample only 

```{r normalize_counts_NOB1,dependson="load_plates"}

# normalized Cq for NOB1 2nd deriviation
plate_cq_data_NOB1_2nd_derivative <- plate_cq_data_NOB1_2nd_derivative %>%
  filter(!prep_type == "-RT")
# the normalisation can use mean or median(,norm_function = median/mean)
plate_norm_PGK1_NOB1_2nd_derivative <- plate_cq_data_NOB1_2nd_derivative %>%
  calculate_deltacq_bysampleid(ref_target_ids = "PGK1")
plate_norm_RPS3_NOB1_2nd_derivative <- plate_cq_data_NOB1_2nd_derivative %>%
  calculate_deltacq_bysampleid(ref_target_ids = "RPS3")
plate_norm_RPS13_NOB1_2nd_derivative <- plate_cq_data_NOB1_2nd_derivative %>%
  calculate_deltacq_bysampleid(ref_target_ids = "RPS13")
plate_norm_mean_NOB1_2nd_derivative <- plate_cq_data_NOB1_2nd_derivative %>%
  calculate_deltacq_bysampleid(ref_target_ids = c("PGK1", "RPS3", "RPS13"), norm_function = mean)
plate_norm_median_NOB1_2nd_derivative <- plate_cq_data_NOB1_2nd_derivative %>%
  calculate_deltacq_bysampleid(ref_target_ids = c("PGK1", "RPS3", "RPS13"))

```


## Plot normalized data, all reps(except -rt+no sample)

```{r combined_three_datafrome,dependson="bind_plates"}
plate_norm_PGK1_all_2nd_derivative <- plate_norm_PGK1_NOB1_2nd_derivative %>%
  filter(target_id == "NOB1") %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))

plate_norm_RPS3_all_2nd_derivative <- plate_norm_RPS3_NOB1_2nd_derivative %>%
  filter(target_id == "NOB1") %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))

plate_norm_RPS13_all_2nd_derivative <- plate_norm_RPS13_NOB1_2nd_derivative %>%
  filter(target_id == "NOB1") %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))

plate_norm_mean_all_2nd_derivative <- plate_norm_mean_NOB1_2nd_derivative %>%
  filter(target_id == "NOB1") %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))

plate_norm_median_all_2nd_derivative <- plate_norm_median_NOB1_2nd_derivative %>%
  filter(target_id == "NOB1") %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))

```

```{r plot_normalized,dependson="median_counts",fig.height=6,fig.width=8}
strain_colors <- c("#330099","#3366FF","#66CCFF","#FF0000","#FF3399","#FF99CC")
# PGK1 as reference gene

# 2nd deriviation approach
ggplot(data = plate_norm_PGK1_all_2nd_derivative) +
  geom_point(aes(x = strain, y = delta_cq, shape = rep_number, colour = clone),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(y = "ΔCq, (log2 relative abundance change)\n relative to PGK1",
    title = "Normalized_Δ_cq_PGK1_2nd_derivative") +
  facet_wrap(~condition,ncol=3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

# RPS3 as reference gene
# 2nd deriviation approach
ggplot(data = plate_norm_RPS3_all_2nd_derivative) +
  geom_point(aes(x = strain, y = delta_cq, shape = rep_number, colour = clone),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(y = "ΔCq (log2 relative abundance change)\n relative to RPS3",
    title = "Normalized_ΔCq_RPS3_2nd_derivative") +
  facet_wrap(~condition,ncol=3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

# RPS13 as reference gene
# 2nd deriviation approach
ggplot(data = plate_norm_RPS13_all_2nd_derivative) +
  geom_point(aes(x = strain, y = delta_cq, shape = rep_number, colour = clone),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(y = "ΔCq (log2 relative abundance change)\n relative to RPS13",
    title = "Normalized_ΔCq_RPS13_2nd_derivative") +
  facet_wrap(~condition,ncol=3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

# mean of reference gene mean value
# 2nd deriviation approach
ggplot(data = plate_norm_mean_all_2nd_derivative) +
  geom_point(aes(x = strain, y = delta_cq, shape = rep_number, colour = clone),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(y = "ΔCq (log2 relative abundance change)\n relative to reference gene",
    title = "Normalized_ΔCq_mean") +
  facet_wrap(~condition,ncol=3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

# median of reference gene mean value
# 2nd deriviation approach
ggplot(data = plate_norm_median_all_2nd_derivative) +
  geom_point(aes(x = strain, y = delta_cq, shape = rep_number, colour = clone),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(y = "ΔCq (log2 relative abundance change)\n relative to reference gene",
    title = "Normalized_ΔCq_median") +
  facet_wrap(~condition,ncol=3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)
```


## Calculate a summary value for each sample-target combination

Having made normalized values, we use `group_by` and `summarize` from the `dplyr` package to calculate the median values of `delta_cq` and `rel_abund` for each sample. This makes a smaller table by collapsing 3 technical replicates into a single summary value for each `sample_id`-`target_id` combination.

```{r median_counts,dependson="load_plates", cache=TRUE}

# median_delta_cq+mean_normalised
plate_med_mean_2nd_derivative <- plate_norm_mean_all_2nd_derivative %>%
  group_by(sample_id, prep_type, target_id) %>%
  summarize(
    delta_cq  = median(delta_cq, na.rm = TRUE),
    rel_abund = median(rel_abund, na.rm = TRUE))
plate_med_mean_2nd_derivative <- plate_med_mean_2nd_derivative %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))

# median_delta_cq+median_normalised
# 2nd derivation
plate_med_median_2nd_derivative <- plate_norm_median_all_2nd_derivative %>%
  group_by(sample_id, prep_type, target_id) %>%
  summarize(
    delta_cq  = median(delta_cq, na.rm = TRUE),
    rel_abund = median(rel_abund, na.rm = TRUE))
plate_med_median_2nd_derivative <- plate_med_median_2nd_derivative %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))

```

## Plot delta Cq

```{r plot_deltacq_vs_gene_total,dependson="median_counts",fig.height=4,fig.width=6}

strain_colors <- c("#330099","#3366FF","#66CCFF","#FF0000","#FF3399","#FF99CC")
ggplot(data = plate_med_mean_2nd_derivative ) + 
  geom_point(aes(x = strain, y = delta_cq, shape = rep_number, colour = clone) ) + 
  labs(
    y = "ΔCq (log2 relative abundance change)\n relative to MEAN of reference gene"
, title = "ΔCq_mean") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  facet_wrap(~condition,ncol=3)+
  scale_color_manual(values = strain_colors)


ggplot(data = plate_med_median_2nd_derivative ) + 
  geom_point(aes(x = strain, y = delta_cq, shape = rep_number, colour = clone) ) + 
  labs(
    y = "ΔCq (log2 relative abundance change)\n relative to MEDIAN of reference gene"
, title = "ΔCq_median") +
  facet_wrap(~condition,ncol=3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

```
## generate deltadelta Cq
```{r generate_deltadeltacq_vs_gene_total,dependson="median_counts",fig.height=6,fig.width=8}
# mean
nonHS <-  plate_norm_mean_all_2nd_derivative %>%
  filter(condition == "nonHS")
nonHS_deltadeltaCq <- nonHS %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_nonHS_rep1","WT_nonHS_rep2"))

shortHS <- plate_norm_mean_all_2nd_derivative %>%
  filter(condition == "shortHS")
shortHS_deltadeltaCq <- shortHS %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_shortHS_rep1","WT_shortHS_rep2"))

mediumHS <- plate_norm_mean_all_2nd_derivative %>%
  filter(condition == "mediumHS")
mediumHS_deltadeltaCq <- mediumHS %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_mediumHS_rep1","WT_mediumHS_rep2"))

deltadelta_Cq_HS_mean <- rbind(nonHS_deltadeltaCq, shortHS_deltadeltaCq, mediumHS_deltadeltaCq)
deltadelta_Cq_HS_mean <- deltadelta_Cq_HS_mean %>%
    group_by(sample_id, prep_type, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE))

deltadelta_Cq_HS_mean <- deltadelta_Cq_HS_mean %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))
deltadelta_Cq_HS_mean

# median
nonHS <-  plate_norm_median_all_2nd_derivative %>%
  filter(condition == "nonHS")
nonHS_deltadeltaCq <- nonHS %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_nonHS_rep1","WT_nonHS_rep2"))

shortHS <- plate_norm_median_all_2nd_derivative %>%
  filter(condition == "shortHS")
shortHS_deltadeltaCq <- shortHS %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_shortHS_rep1","WT_shortHS_rep2"))

mediumHS <- plate_norm_median_all_2nd_derivative %>%
  filter(condition == "mediumHS")
mediumHS_deltadeltaCq <- mediumHS %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_mediumHS_rep1","WT_mediumHS_rep2"))

deltadelta_Cq_HS_median <- rbind(nonHS_deltadeltaCq, shortHS_deltadeltaCq, mediumHS_deltadeltaCq)
deltadelta_Cq_HS_median <- deltadelta_Cq_HS_median %>%
    group_by(sample_id, prep_type, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE))

deltadelta_Cq_HS_median <- deltadelta_Cq_HS_median %>%
  mutate(strain = str_replace(sample_id, "_.*$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"),
         condition = str_extract(sample_id, "(?<=_)[^_]+(?=_rep\\d)"),
         clone = str_extract(sample_id, ".*(?=_rep\\d$)"))
deltadelta_Cq_HS_median

```

## plot deltadeltaCq
```{r plot_deltadeltacq_vs_gene_total,dependson="median_counts",fig.height=4,fig.width=6}

strain_colors <- c("#330099","#3366FF","#66CCFF","#FF0000","#FF3399","#FF99CC")
ggplot(data = deltadelta_Cq_HS_mean ) + 
  geom_point(aes(x = strain, y = deltadelta_cq, shape = rep_number, colour = clone) ) + 
  labs(
    y = "ΔΔCq (log2 fold change)\n relative to WT"
, title = "ΔΔCq_mean_2nd_derivative_condition") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  facet_wrap(~condition,ncol=3)+
  scale_color_manual(values = strain_colors)


ggplot(data = deltadelta_Cq_HS_median ) + 
  geom_point(aes(x = strain, y = deltadelta_cq, shape = rep_number, colour = clone) ) + 
  labs(
    y = "ΔΔCq (log2 fold change)\n relative to WT"
, title = "ΔΔCq_median_2nd_derivative_condition") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  facet_wrap(~condition,ncol=3)+
  scale_color_manual(values = strain_colors)

ggplot(data = deltadelta_Cq_HS_median ) + 
  geom_point(aes(x = condition, y = deltadelta_cq, shape = rep_number, colour = clone) ) + 
  labs(
    y = "ΔΔCq (log2 fold change)\n relative to WT"
, title = "ΔΔCq_median_2nd_derivative_strain") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  facet_wrap(~strain,ncol=2)+
  scale_color_manual(values = strain_colors)

```