---
output:
  html_document: default
---

# Summary: NOB1-FCP1-ARG8 - RT-qPCR experiment v2.0_draft.

## Setup knitr options and load packages

```{r setup,warning=FALSE,message=FALSE}

# knitr options for report generation
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE,
  results = "show")

# Load packages
library(tidyr)
library(ggplot2)
library(dplyr)
library(readr)
library(tidyqpcr)
library(stringr)

# set default theme for graphics
theme_set(theme_bw(base_size = 11) %+replace%
  theme(
    strip.background = element_blank(),
    panel.grid = element_blank()
  ))
```

## Organize sample and target data on the plates

```{r set_platedata_NOB1,dependson="label_plates"}
# NOB1_plate
# set unexclude wells for NOB1
letters_series <- rep(LETTERS[5:16], each = 24)
numbers_series <- rep(1:24, times = 3)
wells_exclude <- paste0(letters_series, numbers_series)

plate_cq_extdata_NOB1 <- read_lightcycler_1colour_cq("WD_24072023_NOB1+ARG8+FCP1_ct.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

plate_cq_extdata_NOB1_2nd_derivative <- read_lightcycler_1colour_cq("WD_24072023_NOB1+ARG8+FCP1_ct_2nd_derivative.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

plate_cq_extdata_NOB1
plate_cq_extdata_NOB1_2nd_derivative

# set plate distribution
plate_cq_data_NOB1 <- 
  create_blank_plate(well_row = LETTERS[1:4], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_NOB1, by = "well") # %>% # join that info to loaded data

plate_cq_data_NOB1_2nd_derivative <- 
  create_blank_plate(well_row = LETTERS[1:4], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_NOB1_2nd_derivative, by = "well") # %>% # join that info to loaded data

target_id_levels <- c("NOB1","PGK1","RPS13","RPS3")
rowkey4 <- tibble(
  well_row = LETTERS[1:4],
  target_id = target_id_levels)
sample_id_levels <- c("WT_rep1", "WT_rep2", "WT_rep3", "NOB1_3'utr_swap_rep1", "NOB1_3'utr_swap_rep2", "NOB1_3'utr_swap_rep3")
prep_type_values <- c(rep("+RT", times = 18), rep("-RT", times = 6))
colkey12 <- tibble(
  well_col = 1:24,
  sample_id = rep(sample_id_levels, times = 4),
  prep_type = prep_type_values)

plate_cq_data_NOB1 <- label_plate_rowcol(
  plate = plate_cq_data_NOB1,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_NOB1_2nd_derivative <- label_plate_rowcol(
  plate = plate_cq_data_NOB1_2nd_derivative,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_NOB1
plate_cq_data_NOB1_2nd_derivative

```

```{r set_platedata_AGR8,dependson="label_plates"}
# ARG8_plate
# set unexclude wells for ARG8
letters_series <- c(rep(LETTERS[1:4], each = 24), rep(LETTERS[9:16], each = 24))
numbers_series <- rep(1:24, times = 3)
wells_exclude <- paste0(letters_series, numbers_series)

plate_cq_extdata_ARG8 <- read_lightcycler_1colour_cq("WD_24072023_NOB1+ARG8+FCP1_ct.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

plate_cq_extdata_ARG8_2nd_derivative <- read_lightcycler_1colour_cq("WD_24072023_NOB1+ARG8+FCP1_ct_2nd_derivative.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

plate_cq_extdata_ARG8
plate_cq_extdata_ARG8_2nd_derivative

# set plate distribution
plate_cq_data_ARG8 <- 
  create_blank_plate(well_row = LETTERS[5:8], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_ARG8, by = "well") # %>% # join that info to loaded data

plate_cq_data_ARG8_2nd_derivative <- 
  create_blank_plate(well_row = LETTERS[5:8], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_ARG8_2nd_derivative, by = "well") # %>% # join that info to loaded data

target_id_levels <- c("ARG8","PGK1","RPS13","RPS3")
rowkey4 <- tibble(
  well_row = LETTERS[5:8],
  target_id = target_id_levels)
sample_id_levels <- c("WT_rep1", "WT_rep2", "WT_rep3", "ARG8_3'utr_swap_rep1", "ARG8_3'utr_swap_rep2", "ARG8_3'utr_swap_rep3")
prep_type_values <- c(rep("+RT", times = 18), rep("-RT", times = 6))
colkey12 <- tibble(
  well_col = 1:24,
  sample_id = rep(sample_id_levels, times = 4),
  prep_type = prep_type_values)

plate_cq_data_ARG8 <- label_plate_rowcol(
  plate = plate_cq_data_ARG8,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_ARG8_2nd_derivative <- label_plate_rowcol(
  plate = plate_cq_data_ARG8_2nd_derivative,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_ARG8
plate_cq_data_ARG8_2nd_derivative

```

```{r set_platedata_FCP1_1,dependson="label_plates"}
# FCP1_plate_1
# set unexclude wells for FCP1
letters_series <- c(rep(LETTERS[1:8], each = 24), rep(LETTERS[13:16], each = 24))
numbers_series <- rep(1:24, times = 3)
wells_exclude <- paste0(letters_series, numbers_series)

plate_cq_extdata_FCP1 <- read_lightcycler_1colour_cq("WD_24072023_NOB1+ARG8+FCP1_ct.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

plate_cq_extdata_FCP1_2nd_derivative <- read_lightcycler_1colour_cq("WD_24072023_NOB1+ARG8+FCP1_ct_2nd_derivative.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

plate_cq_extdata_FCP1
plate_cq_extdata_FCP1_2nd_derivative

# set plate distribution
plate_cq_data_FCP1 <- 
  create_blank_plate(well_row = LETTERS[9:12], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_FCP1, by = "well") # %>% # join that info to loaded data

plate_cq_data_FCP1_2nd_derivative <- 
  create_blank_plate(well_row = LETTERS[9:12], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_FCP1_2nd_derivative, by = "well") # %>% # join that info to loaded data

target_id_levels <- c("FCP1","PGK1","RPS13","RPS3")
rowkey4 <- tibble(
  well_row = LETTERS[9:12],
  target_id = target_id_levels)
sample_id_levels <- c("WT_rep1", "WT_rep2", "WT_rep3", "FCP1_3'utr_swap_A_rep1", "FCP1_3'utr_swap_A_rep2", "FCP1_3'utr_swap_B_rep1")
prep_type_values <- c(rep("+RT", times = 18), rep("-RT", times = 6))
colkey12 <- tibble(
  well_col = 1:24,
  sample_id = rep(sample_id_levels, times = 4),
  prep_type = prep_type_values)

plate_cq_data_FCP1 <- label_plate_rowcol(
  plate = plate_cq_data_FCP1,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_FCP1_2nd_derivative <- label_plate_rowcol(
  plate = plate_cq_data_FCP1_2nd_derivative,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_FCP1
plate_cq_data_FCP1_2nd_derivative

```

```{r set_platedata_FCP1,dependson="label_plates"}
# FCP1_plate_2
# set unexclude wells for FCP1
letters_series <- rep(LETTERS[1:12], each = 24)
numbers_series <- rep(1:24, times = 3)
wells_exclude <- paste0(letters_series, numbers_series)

plate_cq_extdata_FCP1_plus <- read_lightcycler_1colour_cq("WD_24072023_NOB1+ARG8+FCP1_ct.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

plate_cq_extdata_FCP1_2nd_derivative_plus <- read_lightcycler_1colour_cq("WD_24072023_NOB1+ARG8+FCP1_ct_2nd_derivative.txt") %>%
  dplyr::filter(! well %in% wells_exclude ) %>% # exclude blank wells
  dplyr::select(well, sample_info, cq) 

# set plate distribution
plate_cq_data_FCP1_plus <- 
  create_blank_plate(well_row = LETTERS[13:16], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_FCP1_plus, by = "well") # %>% # join that info to loaded data

plate_cq_data_FCP1_2nd_derivative_plus <- 
  create_blank_plate(well_row = LETTERS[13:16], well_col = 1:24) %>% # row and column labels for each well
  inner_join(plate_cq_extdata_FCP1_2nd_derivative_plus, by = "well") # %>% # join that info to loaded data

target_id_levels <- c("FCP1","PGK1","RPS13","RPS3")
rowkey4 <- tibble(
  well_row = LETTERS[13:16],
  target_id = target_id_levels)
sample_id_levels <- c("FCP1_3'utr_swap_B_rep2", "no_sample", "no_sample", "no_sample", "no_sample","no_sample")
prep_type_values <- c(rep("+RT", times = 18), rep("-RT", times = 6))
colkey12 <- tibble(
  well_col = 1:24,
  sample_id = rep(sample_id_levels, times = 4),
  prep_type = prep_type_values)

plate_cq_data_FCP1_plus <- label_plate_rowcol(
  plate = plate_cq_data_FCP1_plus,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_FCP1_2nd_derivative_plus <- label_plate_rowcol(
  plate = plate_cq_data_FCP1_2nd_derivative_plus,
  rowkey = rowkey4,
  colkey = colkey12)

plate_cq_data_FCP1 <- rbind(plate_cq_data_FCP1, plate_cq_data_FCP1_plus)
plate_cq_data_FCP1_2nd_derivative <- rbind(plate_cq_data_FCP1_2nd_derivative, plate_cq_data_FCP1_2nd_derivative_plus)

plate_cq_data_FCP1
plate_cq_data_FCP1_2nd_derivative
```

## Plot unnormalized Cq data

```{r plot_unnormalized_NOB1,dependson="load_plates",fig.height=6,fig.width=8}

plate_cq_data <- rbind(plate_cq_data_ARG8_2nd_derivative, plate_cq_data_FCP1_2nd_derivative, plate_cq_data_NOB1_2nd_derivative)

plate_cq_data <- plate_cq_data %>%
  filter(!sample_id == "no_sample") %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))
# 2nd derivative approach
strain_colors <- c("blue", "red","pink", "green4","black")
ggplot(data = plate_cq_data) +
  geom_point(aes(x = target_id, y = cq, shape = rep_number, colour = strain),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(x= "Target mRNA",
    y = "Quantification cycle (Cq)",
    title = "Unnormalized_NOB1_2nd_derivative_cq"
  ) +
  facet_wrap(~prep_type,ncol=5) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)


```


# Normalized data: delta Cq

## Normalize Cq to PGK1, RPS3 and RPS13, within Sample only 

```{r normalize_counts_NOB1,dependson="load_plates"}
# normalized Cq
plate_cq_data <- plate_cq_data %>%
  filter(!prep_type == "-RT")
# the normalisation can use mean or median(,norm_function = median/mean)
plate_norm_PGK1 <- plate_cq_data %>%
  calculate_deltacq_bysampleid(ref_target_ids = "PGK1")
plate_norm_RPS3 <- plate_cq_data %>%
  calculate_deltacq_bysampleid(ref_target_ids = "RPS3")
plate_norm_RPS13 <- plate_cq_data %>%
  calculate_deltacq_bysampleid(ref_target_ids = "RPS13")
plate_norm_mean <- plate_cq_data %>%
  calculate_deltacq_bysampleid(ref_target_ids = c("PGK1", "RPS3", "RPS13"), norm_function = mean)
plate_norm_median <- plate_cq_data %>%
  calculate_deltacq_bysampleid(ref_target_ids = c("PGK1", "RPS3", "RPS13"))
```

## Plot normalized data, all reps(except -rt+no sample)

```{r combined_three_datafrome,dependson="bind_plates"}
plate_norm_PGK1_all <- plate_norm_PGK1 %>%
  filter(target_id %in% c("NOB1", "ARG8", "FCP1")) %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

plate_norm_RPS3_all <- plate_norm_RPS3 %>%
  filter(target_id %in% c("NOB1", "ARG8", "FCP1")) %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

plate_norm_RPS13_all <- plate_norm_RPS13 %>%
  filter(target_id %in% c("NOB1", "ARG8", "FCP1")) %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

plate_norm_mean_all <- plate_norm_mean %>%
  filter(target_id %in% c("NOB1", "ARG8", "FCP1")) %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

plate_norm_median_all <- plate_norm_median %>%
  filter(target_id %in% c("NOB1", "ARG8", "FCP1")) %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))
```

```{r plot_normalized,dependson="median_counts",fig.height=6,fig.width=8}
strain_colors <- c("blue", "red","pink", "green4","black")
# PGK1 as reference gene
ggplot(data = plate_norm_PGK1_all) +
  geom_point(aes(x = target_id, y = delta_cq, shape = rep_number, colour = strain),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(x= "Target mRNA",y = "ΔCq (log2 relative abundance change)\n relative to PGK1",
    title = "Normalized_ΔCq_PGK1") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

# RPS3 as reference gene
ggplot(data = plate_norm_RPS3_all) +
  geom_point(aes(x = target_id, y = delta_cq, shape = rep_number, colour = strain),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(x= "Target mRNA",y = "ΔCq (log2 relative abundance change)\n relative to RPS3",
    title = "Normalized_ΔCq_RPS3") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

# RPS13 as reference gene
# 2nd deriviation approach
ggplot(data = plate_norm_RPS13_all) +
  geom_point(aes(x = target_id, y = delta_cq, shape = rep_number, colour = strain),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(x= "Target mRNA",y = "ΔCq (log2 relative abundance change)\n relative to RPS13",
    title = "Normalized_ΔCq_RPS13") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

# mean reference gene mean value
# 2nd deriviation approach
ggplot(data = plate_norm_mean_all) +
  geom_point(aes(x = target_id, y = delta_cq, shape = rep_number, colour = strain),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(x= "Target mRNA",y = "ΔCq, (log2 relative abundance change)\n relative to MEAN of reference gene",
    title = "Normalized_ΔCq_mean") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

# 2nd deriviation approach
ggplot(data = plate_norm_median_all) +
  geom_point(aes(x = target_id, y = delta_cq, shape = rep_number, colour = strain),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(x= "Target mRNA",y = "ΔCq, (log2 relative abundance change)\n relative to MEDIAN of reference gene",
    title = "Normalized_ΔCq_median") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)
```


## Calculate a summary value for each sample-target combination

Having made normalized values, we use `group_by` and `summarize` from the `dplyr` package to calculate the median values of `delta_cq` and `rel_abund` for each sample. This makes a smaller table by collapsing 3 technical replicates into a single summary value for each `sample_id`-`target_id` combination.

```{r median_counts,dependson="load_plates"}
# mean
plate_med_mean <- plate_norm_mean_all %>%
  group_by(sample_id, prep_type, target_id) %>%
  summarize(
    delta_cq  = median(delta_cq, na.rm = TRUE),
    rel_abund = median(rel_abund, na.rm = TRUE))
plate_med_mean <- plate_med_mean %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

# median
plate_med_median <- plate_norm_median_all %>%
  group_by(sample_id, prep_type, target_id) %>%
  summarize(
    delta_cq  = median(delta_cq, na.rm = TRUE),
    rel_abund = median(rel_abund, na.rm = TRUE))
plate_med_median <- plate_med_median %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))
```

## Plot delta Cq

```{r plot_deltacq_vs_gene_total,dependson="median_counts",fig.height=6,fig.width=8}
strain_colors <- c("blue", "red","pink", "green4","black")
ggplot(data = plate_med_mean ) + 
  geom_point(aes(x = target_id, y = delta_cq, shape = rep_number, colour = strain) ) + 
  labs(x= "Target mRNA",
    y = "ΔCq (log2 relative abundance change)\n relative to MEAN of reference gene"
, title = "ΔCq_mean") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  scale_color_manual(values = strain_colors)

ggplot(data = plate_med_median ) + 
  geom_point(aes(x = target_id, y = delta_cq, shape = rep_number, colour = strain) ) + 
  labs(x= "Target mRNA",
    y = "ΔCq (log2 relative abundance change)\n relative to MEDIAN of reference gene"
, title = "ΔCq_median") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  scale_color_manual(values = strain_colors)
```

## Calculate deltadelta Cq

```{r calculate_deltadeltacq,dependson="normalize_counts"}
# also can choose median or mean, here is median
plate_delta_mean_norm_mean <- plate_norm_mean_all %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_rep1","WT_rep2","WT_rep3",
                                 norm_function = mean))
plate_delta_mean_norm_mean <- plate_delta_mean_norm_mean %>%
    group_by(sample_id, prep_type, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE))
plate_delta_mean_norm_mean <- plate_delta_mean_norm_mean %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

plate_delta_median_norm_mean <- plate_norm_mean_all %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_rep1","WT_rep2","WT_rep3"))
plate_delta_median_norm_mean <- plate_delta_median_norm_mean %>%
    group_by(sample_id, prep_type, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE))
plate_delta_median_norm_mean <- plate_delta_median_norm_mean %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

plate_delta_mean_norm_median <- plate_norm_median_all %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_rep1","WT_rep2","WT_rep3"),
                                 norm_function = mean)
plate_delta_mean_norm_median <- plate_delta_mean_norm_median %>%
    group_by(sample_id, prep_type, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE))
plate_delta_mean_norm_median <- plate_delta_mean_norm_median %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

plate_delta_median_norm_median <- plate_norm_median_all %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("WT_rep1","WT_rep2","WT_rep3"))
plate_delta_median_norm_median <- plate_delta_median_norm_median %>%
    group_by(sample_id, prep_type, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE))
plate_delta_median_norm_median <- plate_delta_median_norm_median %>%
  mutate(strain = str_replace(sample_id, "_rep\\d$", ""),
         rep_number = str_extract(sample_id, "rep\\d$"))

```

## Plot deltadelta Cq

```{r plot_deltadeltacq_bytarget_mean,dependson="calculate_deltadeltacq",fig.height=4,fig.width=6}
strain_colors <- c("blue", "red","pink", "green4","black")
ggplot(data = plate_delta_mean_norm_mean ) + 
  geom_point(aes(x = target_id, y = deltadelta_cq, colour = strain, shape=rep_number)) + 
  labs(x= "Target mRNA",
    y = "ΔΔCq (log2 fold change)\n relative to WT", title = "ΔΔCq_norm_mean_WT_mean"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

ggplot(data = plate_delta_median_norm_mean ) + 
  geom_point(aes(x = target_id, y = deltadelta_cq, colour = strain, shape=rep_number)) + 
  labs(x= "Target mRNA",
    y = "ΔΔCq (log2 fold change)\n relative to WT", title = "ΔΔCq_norm_mean_wt_median"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

plate_delta_mean_norm_median
ggplot(data = plate_delta_mean_norm_median ) + 
  geom_point(aes(x = target_id, y = deltadelta_cq, colour = strain, shape=rep_number)) + 
  labs(x= "Target mRNA",
    y = "ΔΔCq (log2 fold change)\n relative to WT", title = "ΔΔCq_norm_median_wt_mean"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)

ggplot(data = plate_delta_median_norm_median ) + 
  geom_point(aes(x = target_id, y = deltadelta_cq, colour = strain, shape=rep_number)) + 
  labs(x= "Target mRNA",
    y = "ΔΔCq (log2 fold change)\n relative to WT", title = "ΔΔCq_norm_median_wt_median"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values = strain_colors)
```
